### 2017-03-24 grlp.tex
cleaning todos

While solving the ALP can be significantly cheaper than solving the LP underlying the MDP
and thus it can be advantageous for moderate-scale MDPs,
 \todoc{Give the big-Oh costs!} the number of constraints in the ALP is $SA$, 
hence the ALP is still intractable for huge-scale MDPs.
%While evaluating the objective and checking a single constraint for constraint violations can be done in constant time,
%since the number of constraints in the ALP is equal to the number of state-action pairs, 
%the ALP is still intractable in general.
 To reduce the number of constraints, we consider a relaxation of ALP 
where the constraints are replaced with positive linear 
combinations of them. 
Recalling that the constraints took the form $J \ge g_a + \alpha P_a J$ (with $J = \Phi r$),
choosing $m$ to be target number of constraints, for $1\le i \le m$, the $i$th new constraint is given by
$\sum_a w_{i,a}^\top J \ge \sum_a w_{i,a}^\top(g_a + \alpha P_a J)$, 
where the choice of $m$ and that of the vectors $w_{i,a}\in \Re_+^S$ is left to the user.
Note that this results in a linear program with $k$ variables and $m$ constraints, which can be written as
\begin{align}\label{grlp}
\begin{split}
&\underset{r\in \Re^k}{\min}\, \, c^\top \Phi r\\
&\text{s.t.}\mb  \,\sum_a W_a^\top \Phi r\geq \sum_a W_a^\top (g_a + \alpha P_a) \Phi r\,,
\end{split}
\end{align}
where $W_a = (w_{1,a},\dots,w_{m,a}) \in \Re_+^{S \times m}$. 
\todoc{Do we need that $(w_{i,a}^\top)_i$ defines a probability distribution?}
Note that the $(i,j)$th entry of the $m\times k$ constraint matrix of the resulting LP is
$\sum_a  w_{i,a}^\top  \phi_j - \alpha \sum_a w_{i,a}^\top P_a \phi_j$ and assuming that $(w_{i,a})_{a}$ has $p$ nonzero
elements, this can be calculated in $O( p )$ time, making the total cost of obtaining the constraint matrix to be $O(mkp)$ regardless the value of $S$ and $A$. \todoc{Somewhere discuss that the cost of obtaining a policy will depend on $A$. 
We can also discuss how $A$ could be folded into states.}

We will call the LP in \eqref{grlp} the \emph{linearly relaxed approximate linear program (LRALP)}. \todoc{Replace GRLP everywhere with LRALP.} Any LP obtained using 
any constraint selection/generation process can be represented by choosing an appropriate binary-valued matrix
$W^\top = (W_1^\top,\dots,W_A^\top)\in \Re_+^{m\times SA}$. In particular, when the constraints are selected
in a random process as suggested by \citet{CS}, the matrix $W$ would be a random, binary-valued matrix.

>>>>

While solving the ALP can be significantly cheaper than solving the LP underlying the MDP
and thus it can be advantageous for moderate-scale MDPs,
 \todoc{Give the big-Oh costs!} the number of constraints in the ALP is $SA$, 
hence the ALP is still intractable for huge-scale MDPs.
%While evaluating the objective and checking a single constraint for constraint violations can be done in constant time,
%since the number of constraints in the ALP is equal to the number of state-action pairs, 
%the ALP is still intractable in general.
 To reduce the number of constraints, we consider a relaxation of ALP 
where the constraints are replaced with positive linear 
combinations of them. 
Recalling that the constraints took the form $J \ge g_a + \alpha P_a J$ (with $J = \Phi r$),
choosing $m$ to be target number of constraints, for $1\le i \le m$, the $i$th new constraint is given by
$\sum_a w_{i,a}^\top J \ge \sum_a w_{i,a}^\top(g_a + \alpha P_a J)$, 
where the choice of $m$ and that of the vectors $w_{i,a}\in \Re_+^S$ is left to the user.
Note that this results in a linear program with $k$ variables and $m$ constraints, which can be written as
\begin{align}\label{grlp}
\begin{split}
&\underset{r\in \Re^k}{\min}\, \, c^\top \Phi r\\
&\text{s.t.}\mb  \,\sum_a W_a^\top \Phi r\geq \sum_a W_a^\top (g_a + \alpha P_a) \Phi r\,,
\end{split}
\end{align}
where $W_a = (w_{1,a},\dots,w_{m,a}) \in \Re_+^{S \times m}$. 
Note that the $(i,j)$th entry of the $m\times k$ constraint matrix of the resulting LP is
$\sum_a  w_{i,a}^\top  \phi_j - \alpha \sum_a w_{i,a}^\top P_a \phi_j$ and assuming that $(w_{i,a})_{a}$ has $p$ nonzero
elements, this can be calculated in $O( p )$ time, making the total cost of obtaining the constraint matrix to be $O(mkp)$ regardless the value of $S$ and $A$. \todoc{Somewhere discuss that the cost of obtaining a policy will depend on $A$. 
We can also discuss how $A$ could be folded into states.}

We will call the LP in \eqref{grlp} the \emph{linearly relaxed approximate linear program (LRALP)}. 
Any LP obtained using 
any constraint selection/generation process can be represented by choosing an appropriate binary-valued matrix
$W^\top = (W_1^\top,\dots,W_A^\top)\in \Re_+^{m\times SA}$. In particular, when the constraints are selected
in a random process as suggested by \citet{CS}, the matrix $W$ would be a random, binary-valued matrix.

### 2017-03-24 intro.tex
cleaning todos

The main objective of \emph{planning} is to compute actions of an optimal policy while interacting with an MDP model. 
In finite state-action MDPs,
assuming access to individual transition probabilities and rewards along transitions, 
various algorithms are available to perform this computation in time and space that scales polynomially with the number of states and actions.
However, in most practical applications, the MDP is compactly represented
and if it is not infinite, the number of states scale \emph{exponentially} with the \emph{size of the representation} of the MDP.
%an effect that is known as \emph{Bellman's curse of dimensionality}.
If planners are allowed to perform some fixed amount of calculations for each state encountered,
it is possible to use sampling to make the per-state calculation-cost 
independent of the size of the state space \cite{rust96:randomization,szepesvari2001,kearns2002sparse}.
Nevertheless, the resulting methods are still quite limited. 
In fact, various hardness results show that computing actions of (near-) optimal policies is intractable in various senses
and in various compactly represented MDPs \cite{BlonTsi:00Complexity}.
Given these negative results, 
it is customary to adopt the \emph{modest goal of efficiently computing actions of a policy that 
is nearly as good as a policy chosen by a suitable (computationally unbounded, and well-informed) oracle
from a given restricted policy class}. \todoc{Is this better now? We should probably elaborate on the oracle idea space permitting.}
Here, within some restrictions (see below), the policy class can be chosen by the user.
The more flexibility the user is given in this choice, the stronger a planning method is.

>>>>

The main objective of \emph{planning} is to compute actions of an optimal policy while interacting with an MDP model. 
In finite state-action MDPs,
assuming access to individual transition probabilities and rewards along transitions, 
various algorithms are available to perform this computation in time and space that scales polynomially with the number of states and actions.
However, in most practical applications, the MDP is compactly represented
and if it is not infinite, the number of states scale \emph{exponentially} with the \emph{size of the representation} of the MDP.
%an effect that is known as \emph{Bellman's curse of dimensionality}.
If planners are allowed to perform some fixed amount of calculations for each state encountered,
it is possible to use sampling to make the per-state calculation-cost 
independent of the size of the state space \cite{rust96:randomization,szepesvari2001,kearns2002sparse}.
Nevertheless, the resulting methods are still quite limited. 
In fact, various hardness results show that computing actions of (near-) optimal policies is intractable in various senses
and in various compactly represented MDPs \cite{BlonTsi:00Complexity}.
Given these negative results, 
it is customary to adopt the \emph{modest goal of efficiently computing actions of a policy that 
is nearly as good as a policy chosen by a suitable (computationally unbounded, and well-informed) oracle
from a given restricted policy class}. \todoc{We should probably elaborate on the oracle idea space permitting.
Or in the future. Btw, the situation is similar to statistical learning theory: There competing with the best choice 
is only information theoretically possible. And competing with best choice
in hindsight is often computationally intractable. Maybe cite information theory result for competing with the best policy in a class?}
Here, within some restrictions (see below), the policy class can be chosen by the user.
The more flexibility the user is given in this choice, the stronger a planning method is. \todoc{Add reference to the intro to Mausam's book. But where and what can be said?}

### 2017-03-24 num.tex
cleaning todos

\textbf{Experimental Methodology:} We compare two different sampling strategies $(i)$ based on the \emph{cone} conditions, and $(ii)$ based on constraint sampling. The two strategies are compared via  \emph{lookahead} policies, wherein, the action at state $s$ is obtained by computing the approximate value functions of the next states. The details are as follows:
\emph{Case (i)}: Except for the corner states i.e., $s=0$ and $s=999$, each state $0<s<S-1$ has two next states namely $s'=s-1$ and $s'=s+1$. We formulate two separate LRALPs (or just one LRALP for $s=0$ and $s=S-1$) for next states. When formulating the LRALP for state $s'$, we let $c=e_{s'}$ and choose the constraint corresponding to state $s'$ to ensure the cone condition to be met for LRALP. We choose $5$ more constraints corresponding to states $1,200,400,600,800,999$ (uniformly spaced across the state space) and compute $\hat{J}_{e_{s'}}$. The lookahead policy is formulated as
$u_{LRA}(s)=\argmin_{a\in A} g_a(s)+\sum_{s'\in S}p_a(s,s')\hat{J}_{e_s'}(s')$.

\emph{Case (ii)}: In a manner similar to \emph{Case (i)}, here too we formulate two separate LRALPs for next states. When formulating the LRALP for state $s'$, we first let $c_{s'}(i)=(1-\alpha)(\alpha)^{|s'-i|}$ and normalize $c$ and sample $m=6$ constraints randomly using $c$ and compute $\hat{J}_{c_s'}$. 
The lookahead policy is formulated as 
$u_{CS}(s)=\argmin_{a\in A} g_a(s)+\sum_{s'\in S}p_a(s,s')\hat{J}_{c_{s'}}(s')$.

>>>>

\textbf{Experimental Methodology:} We compare two different sampling strategies $(i)$ based on the \emph{cone} conditions, and $(ii)$ based on constraint sampling. The two strategies are compared via  \emph{lookahead} policies, wherein, the action at state $s$ is obtained by computing the approximate value functions of the next states. The details are as follows:
\emph{Case (i)}: Except for the corner states i.e., $s=0$ and $s=999$, each state $0<s<S-1$ has two next states namely $s'=s-1$ and $s'=s+1$. We formulate two separate LRALPs (or just one LRALP for $s=0$ and $s=S-1$) for next states. When formulating the LRALP for state $s'$, we let $c=e_{s'}$ and choose the constraint corresponding to state $s'$ to ensure the cone condition to be met for LRALP. We choose $5$ more constraints corresponding to states $1,200,400,600,800,999$ (uniformly spaced across the state space) and compute $\hat{J}_{e_{s'}}$. The lookahead policy is formulated as
$u_{LRA}(s)=\argmin_{a\in A} g_a(s)+\sum_{s'\in S}p_a(s,s')\hat{J}_{e_{s'}}(s')$.

\emph{Case (ii)}: In a manner similar to \emph{Case (i)}, here too we formulate two separate LRALPs for next states. When formulating the LRALP for state $s'$, we first let $c_{s'}(s)=(1-\alpha)(\alpha)^{|s'-s|}$ and normalize $c$ and sample $m=6$ constraints randomly using $c$ and compute $\hat{J}_{c_s'}$. 
The lookahead policy is formulated as 
$u_{CS}(s)=\argmin_{a\in A} g_a(s)+\sum_{s'\in S}p_a(s,s')\hat{J}_{c_{s'}}(s')$.
